### 2015-01-31 doc/thesis/evaluation.tex
Minor improvements to the text about related work

We performed some case studies on existing Lua libraries
to evaluate the applicability of our type system.
For each library, we used Typed Lua to either rewrite its modules
or to write statically typed interfaces to its modules through
Typed Lua's description files.
In Section \ref{sec:cases} we present our evaluation results
and discuss some interesting cases.

We also compared our type system to other type systems,
as a way to evaluate our contributions.
In Section \ref{sec:relatedwork} we discuss what are the novel
features introduced by our type system,
what are the features that it shares with other type systems,
and whether we can incorporate some interesting features from
other type systems.

@@

We selected seven case studies to evaluate two aspects of Typed Lua:
\begin{enumerate}
\item how precisely it can describe the type of the interface of a module;
\item if it provides guarantees that the code corresponds to the interface.
\end{enumerate}

@@

The second case study that we chose was the MD5 library \citep{lmd5},
because it is simple and contains just one module.
We needed a simple case study to introduce Typed Lua's description
files and \texttt{userdata} declarations.
These Typed Lua's mechanisms allow programmers to give statically typed
interfaces to Lua libraries that are written in C.

@@

There are three case studies that we randomly selected from the LuaRocks
repository, they are: HTTP Digest \citep{luahttpdigest},
Typical \citep{luatypical}, and Mod 11 \citep{luamod11}.
The first one provides client side HTTP digest authentication for Lua.
The second one is an extension to the primitive function \texttt{type}.
The third one is a generator and checker of modulo 11 numbers.
We randomly selected these libraries to evaluate Typed Lua for
rewriting existing Lua modules.

The Typed Lua compiler is the last case study that we evaluated.
We chose it as a case study because it is a large application.
In addition, bootstrapping the compiler also provides some
confidence on the type system.

@@

In this section we compare Typed Lua with other projects.
We split this section into two parts: the first compares Typed Lua
with other Lua projects, while the second compares Typed Lua with
other projects that are not related to Lua.

@@

MoonScript \citep{moonscript} is a programming language that supports
class-based object-oriented programming, and compiles to idiomatic
Lua code.
However, it does not perform compile-time type checking.

LuaInspect \citep{luainspect} is a tool that uses MetaLua to perform
some code analysis, such as flagging unknown global variables and
table fields, checking the number of function arguments against
signatures, and inferring function return values, but it does not
try to analyze object-oriented code and does not perform compile-time
type checking.

Tidal Lock \citep{tidallock} is a prototype of another optional type
system for Lua, which is written in Metalua.
Tidal Lock covers a little subset of Lua.
Statements include declaration of local variables, multiple assignments,
function application, and the return statement, that is, they do not
include any control-flow statements.
Expressions include primitive literals, table indexing, function application,
function declaration, and the table constructor, but they do not include
binary operations.

@@

Tidal Lock is also a structural type system based on subtyping that has
local type inference.
However, it does not support union types, recursive types, and variadic types.
It also does not type any object-oriented idiom.

@@

@@

Dart \citep{dart} is a new class-based object-oriented programming
language.
It includes optional type annotations and compiles to JavaScript.
The type system of Dart is nominal and includes base types,
function types, lists, and maps.
It also supports generics, and the programmer can define
generic functions, lists, and maps.

@@

TypeScript \citep{typescript} is a JavaScript extension
that includes optional type annotations and class-based
object-oriented programming.
It also uses local type inference to deduce the type
of unannotated expressions.
The type system of TypeScript is structural, based
on subtyping, and supports generics.
It includes the dynamic type, primitive types, function
types, array types, tuple types, recursive types, and
object types.
Unlike Typed Lua, TypeScript does not include union types,
and it uses arrays to represent variadic functions and
multiple return values.

>>>>

We performed some case studies on existing Lua libraries
to evaluate the applicability of our type system.
For each library, we used Typed Lua to either annotate its modules
or to write statically typed interfaces to its modules through
Typed Lua's description files.
In Section \ref{sec:cases} we present our evaluation results
and discuss some interesting cases.

We also evaluate our contributions through the revision of related work.
In Section \ref{sec:relatedwork} we discuss what are the novel
features of our type system, what are the features that it shares
with other type systems, and whether we can incorporate some
interesting features from other type systems.

@@

We selected seven case studies to evaluate two aspects of Typed Lua:
\begin{enumerate}
\item how precisely it can describe the type of the interface of a module;
\item whether it provides guarantees that the code matches the interface.
\end{enumerate}

@@

The second case study that we chose was the MD5 library \citep{lmd5},
because we wanted a simple case study to introduce Typed Lua's description
files and \texttt{userdata} declarations.
These Typed Lua's mechanisms allow programmers to give statically typed
interfaces to Lua libraries.

@@

There are three case studies that we randomly selected from the LuaRocks
repository, they are: HTTP Digest \citep{luahttpdigest},
Typical \citep{luatypical}, and Mod 11 \citep{luamod11}.
The first one provides client side HTTP digest authentication for Lua.
The second one is an extension to the primitive function \texttt{type}.
The third one is a generator and checker of modulo 11 numbers.
We randomly selected these libraries to evaluate Typed Lua for
annotating existing Lua modules.

The Typed Lua compiler is the last case study that we evaluated.
We chose the Typed Lua compiler as a case study because it is a large application,
and bootstrapping the compiler also provides some confidence in the type system.

@@

In this section we review related work.
We split this section into two parts: in the first part we review
other Lua projects, while in the second part we review other
projects that are not related to Lua.

@@

MoonScript \citep{moonscript} is a programming language that supports
class-based object-oriented programming.
MoonScript compiles to idiomatic Lua code, but 
it does not perform compile-time type checking.

LuaInspect \citep{luainspect} is a tool that uses MetaLua to perform
some code analysis.
For instance, it flags unknown global variables and table fields,
it checks the number of function arguments against signatures, and
it infers function return values.
However, it does not try to analyze object-oriented code and
it does not perform compile-time type checking.

Tidal Lock \citep{tidallock} is a prototype of another optional type
system for Lua, which is written in Metalua.
Tidal Lock covers a little subset of Lua.
Statements include declaration of local variables, multiple assignments,
function application, and the return statement.
This means that Tidal Lock does not include any control-flow statement.
Expressions include primitive literals, table indexing, function application,
function declaration, and the table constructor, but they do not include
binary operations.

@@

Tidal Lock is also a structural type system that relies on subtyping and
local type inference.
However, it does not support union types, recursive types, and variadic types.
It also does not type any object-oriented idiom.

@@

Ravi \citep{ravi} is an experimental Lua dialect.
Ravi introduces optional static typing for Lua to improve run-time performance.
To do that, Ravi extends the Lua Virtual Machine to include new
operations that take into account static type information.

@@

Dart \citep{dart} is a new class-based object-oriented programming
language.
It includes optional type annotations and compiles to JavaScript.
The type system of Dart is nominal and includes base types,
function types, lists, and maps.
It also supports generics, and the programmer can define
generic functions, lists, and maps.
Unlike Typed Lua, Dart is unsound by design.

@@

TypeScript \citep{typescript} is a JavaScript extension
that includes optional type annotations and class-based
object-oriented programming.
It also uses local type inference to deduce the type
of unannotated expressions.
The type system of TypeScript is structural, based
on subtyping, and supports generics.
It includes the dynamic type, primitive types, union types,
function types, array types, tuple types, recursive types, and
object types.
Unlike Typed Lua, TypeScript uses arrays to represent variadic
functions and multiple return values.
